---
title: "Dashboard Pauline"
author: "Pauline Sayn-Wittgenstein" 
date: "February-March 2020"

output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    theme: flatly 
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(spotifyr)
library(compmus)
source('spotify.R')
```


The Role of Timbre 
=====================================  

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Cepstogram - Timbre: German Hip Hop outlier - Wenn der Vorhang fällt (feat.Wasi) by Freundeskreis

```{r}
WennderV <- 
     get_tidy_audio_analysis('5GEPmgYRf9e1hcnsbfG9Rz') %>% 
     compmus_align(bars, segments) %>% 
     select(bars) %>% unnest(bars) %>% 
     mutate(
         pitches = 
             map(segments, 
                 compmus_summarise, pitches, 
                 method = 'rms', norm = 'euclidean')) %>% 
     mutate(
         timbre = 
             map(segments, 
                 compmus_summarise, timbre, 
                 method = 'rms', norm = 'euclidean'))
WennderV %>% 
     compmus_gather_timbre %>% 
     ggplot(
         aes(
             x = start + duration / 2, 
             width = duration, 
             y = basis, 
             fill = value)) + 
     geom_tile() +
     labs(x = 'Time (s)', y = NULL, fill = 'Magnitude') +
     scale_fill_viridis_c(option = 'E') +
     theme_classic()
```

### Cepstogram - Timbre: Reggeaton outlier - No Le Temas a el Beat by Trebol Clan and Hector & Tito 

```{r}
Noletemas <- 
     get_tidy_audio_analysis('3rMyZL6J4t2EeZKMTD1UaI') %>% 
     compmus_align(bars, segments) %>% 
     select(bars) %>% unnest(bars) %>% 
     mutate(
         pitches = 
             map(segments, 
                 compmus_summarise, pitches, 
                 method = 'rms', norm = 'euclidean')) %>% 
     mutate(
         timbre = 
             map(segments, 
                 compmus_summarise, timbre, 
                 method = 'rms', norm = 'euclidean'))
Noletemas %>% 
    compmus_gather_timbre %>% 
    ggplot(
        aes(
            x = start + duration / 2, 
            width = duration, 
            y = basis, 
            fill = value)) + 
    geom_tile() +
    labs(x = 'Time (s)', y = NULL, fill = 'Magnitude') +
    scale_fill_viridis_c(option = 'E') +
    theme_classic()
```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Self-Similarity Matrix - Wenn der Vorhang fällt (feat.Wasi) by Freundeskreis

```{r}
WennderV %>% 
    compmus_self_similarity(timbre, 'euclidean') %>% 
    ggplot(
        aes(
            x = xstart + xduration / 2, 
            width = xduration,
            y = ystart + yduration / 2,
            height = yduration,
            fill = d)) + 
    geom_tile() +
    coord_fixed() +
    scale_fill_viridis_c(option = 'E', guide = 'none') +
    theme_classic() +
    labs(x = '', y = '')
```

### Self-Similarity Matrix - No Le Temas a el Beat by Trebol Clan and Hector & Tito 

```{r}
Noletemas %>% 
    compmus_self_similarity(timbre, 'manhattan') %>% 
    ggplot(
        aes(
            x = xstart + xduration / 2, 
            width = xduration,
            y = ystart + yduration / 2,
            height = yduration,
            fill = d)) + 
    geom_tile() +
    coord_fixed() +
    scale_fill_viridis_c(option = 'E', guide = 'none') +
    theme_classic() +
    labs(x = '', y = '')
```

Row 
-----------------------------------------------------------------------
For the cepstograms and the self-similarity matrices I chose the two outliers of the two playlists.

When comparing the cepstograms regarding timbre we can see that the magnitude for the Reggeaton song is much less distributed than the magnitude for the German song, although the German song highlights c03. As these two are outliers for the actual genres we can again see that variety in tone and in timbre, as it is not the case for the Reggeaton outlier, might indeed a characteristic for Reggeaton. To be sure of that, I want to make further investigations on that and possibly also create cepstograms for typical Reggeaton songs. 

The self-similarity matrices that also portray the two outliers are very diverse as well. Although these two are outliers for each genre we can see that there is a clear difference between the two. To find a more clear and exact interpretation for these, I would like to create more self-similarity matrices with different songs from each playlist. 

Popularity of Reggeaton  
===================================== 

Column {data-width=500}
-----------------------------------------------------------------------

### Introduction 

I decided to focus on the difference between Reggeaton and German Hip Hop for my research. Therefore my final research question is:

How does the latin Hip Hop music style "Reggeaton" differ from German Hip Hop? 
Why is Reggeaton growing in popularity all around the world and German Hip Hop is not? Can we find specific features in Reggeaton music that might possibly cause its growing popularity? 

My first motivation for this research is my personal interest in both music styles. Additionally, I find it very impressive that Reggeaton's popularity is growing so rapidly. I personally know a lot of people who do not speak any Spanish but love to listen to Reggeaton music. Therefore, I find it very intersting to find out about that reason that makes people like Reggeaton so much.

Because after doing several analyses with different playlists created by Spotify and getting good suggestions from peers, I realized, it might be better to use playlists that include new but also old songs and not only the current Charts. I also decided not to use the playlists I created but rather use the most accurate ones I can find created by Spotify as this excludes my musical taste. Therefore, I do not project the tables and data from the analyses in this assignment but only the ones I created afterwards with the new playlists. 

The playlist I am using for the music style of Reggeaton is called "Reggeaton Classics" and includes 99 songs.
For the German Hip Hop I found a playlist by Spotify called "Deutschrap: Die Klassiker" but it only includes 50 songs. As I did not know if that is a problem I decided to create a new playlist including this and a second playlist by Spotify called "Modus Mio", and according to Spotify, it is the most important Hip-Hop playlists in Germany. I called this new playlist "German HipHop Classics". This playlist therefore ended with 100 songs. 

When looking at the means and standard deviations of both these playlists I could find small differences between the danceability, loudness and speechiness. 

The small differences that I could find from my first anaylses also reflect my expectations resulting from my own musicial experiences with both musical styles. I myself experience Reggeaton as very moving, danceable and often positive while German Hip Hop is often more negative and aggresive which might result in its loudness. Also, it seems to me that in the German songs I perceive much more text that is sung in a very fast way. Additionally, I perceive a very characteristic beat in most Reggeaton songs.
But of course there are also variations within each musical style and thats what I also saw when doing my first anaylses.


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Visualisation 1

```{r}

Reggeaton_username <- 'spotify'
German_username <- 'paulinchen1999'
Reggeaton_uris <- '37i9dQZF1DX8SfyqmSFDwe'
German_uris <- '3Q6aTzn71rrODobApwJlWR'

Reggeaton <- get_playlist_audio_features(Reggeaton_username, Reggeaton_uris)
German <- get_playlist_audio_features(German_username, German_uris)
both <- Reggeaton %>% mutate(playlist = "Reggeaton Classics") %>%bind_rows(German %>% mutate(playlist = "German HipHop Classics"))

genres <- both %>% mutate(mode=factor(mode,c(1,0),c("Major","Minor")))%>% ggplot(aes(x=valence, y=energy, size=loudness, colour=mode, label=track.name))+geom_point()+geom_rug(size=0.1)+facet_wrap(~playlist)+scale_x_continuous(limits=c(0,1),breaks=c(0, 0.50, 1),  minor_breaks = NULL) +scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.50, 1),minor_breaks = NULL) +scale_colour_brewer(type="qual",palette="Paired")+scale_size_continuous(trans="exp",guide="none")+theme_light()+labs(x="Valence", y="Energy", colour="Mode")

ggplotly(genres)


``` 

### Interpretation 

Here is the first visualisation I created for my research. It compares the valence on the x-axis to the energy of the song on the y-axis.From the first impression you might not really see a big difference between the German and the Reggeaton songs. But looking closer, you can see that there is more variance within the German songs than within the Reggeaton songs. As I described in my own expectations, Reggeaton songs sound very positive and as the visualisation shows, German songs vary regarding valence while the Reggeaton songs almost all score very high on it. What we can also see is that both genres are quite energetic although, I think, this energy might be different for both of them. The energy in German songs seems more aggresive to me than the energy in the Reggeaton songs. This might also be visualised in the graph, as the German songs are also destributed more central than the Reggeaton songs.
Overall from this result, I would describe Reggeaton songs as more positive and more energetic than German Hip Hop songs. This positivity and energy might also be one of the reasons why Reggeaton is so popular among many people all aroung the world. Its positive and energetic vibe spreads a good mood and the language of the songs do not seem to matter in this regard. German Hip Hop songs on the otherhand might also be very energetic but in different way. Looking at the speechiness in the songs, which i perceive to be very high in German songs could be an interesting factor. 

```{r}


```
Variation - The Key to Success? 
=====================================     

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Chromagram - German song "Mios mit Bars" by Luciano

```{r}
Mios <- get_tidy_audio_analysis('7Ek9e3eIuktIFjpDRQfmHE') %>% select(segments) %>% unnest(segments) %>%  select(start, duration, pitches)

Mios %>% 
     mutate(pitches = map(pitches, compmus_normalise, 'euclidean')) %>% 
     compmus_gather_chroma %>% 
     ggplot(
         aes(
             x = start + duration / 2, 
             width = duration, 
             y = pitch_class, 
             fill = value)) + 
     geom_tile() +
     labs(x = 'Time (s)', y = NULL, fill = 'Magnitude') +
     theme_minimal()

```

### Chromagram - Reggeaton song "En la Cama" by Nicky Jam feat. Daddy Yankee

```{r}
Enlacama<- get_tidy_audio_analysis('2Eg6dOam7cAe5turf2bnCg') %>% 
     select(segments) %>% unnest(segments) %>% 
     select(start, duration, pitches)

Enlacama %>%  mutate(pitches = map(pitches, compmus_normalise, 'euclidean')) %>% 
     compmus_gather_chroma %>% 
     ggplot(
         aes(
             x = start + duration / 2, 
             width = duration, 
             y = pitch_class, 
             fill = value)) + 
     geom_tile() +
     labs(x = 'Time (s)', y = NULL, fill = 'Magnitude') +
     theme_minimal()

```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------
I created these two Chromagrams choosing the first song on each playlist. We can see that the first song "Mios mit Bars" by Luciano from the German Hip Hop Playlist is mostly played in A and not that much in G#, G and F#. Additionally, we can see that in the beginning and the end of the measured time of the Chromagram for a lot of notes the magnitude is very low. When looking at the second Chromagram visualising the song "En la Cama" by Nicky Jam featuring Daddy Yankee we can see that the magnitude is higher for almost all notes. Also, the ending and beginning measured for this song's chromgram do not exhibit such striking lower magnitude as the German song. 
From these visualisations we might therefore again conclude that this variety of notes and therefore sounds in the Reggeaton song might lead to more popularity among listeners from all around the world than the more monotony of the German song.  


First conclusions 
=====================================  
    
### Conclusion

What I can conclude from my visualisations until now, is that there might be significant differences between the two music styles, for instance in their level of energy and valence which might be a reason for why Reggeaton is also listened to by people who do not even speak Spanish. It might be the positive, danceable and energetic sound that makes people from all around the world enjoy listening Reggeaton.

I also can conlcude now that from my visualisations I can perceive more variability in tone and timbre for Reggeaton songs than for German Hip Hop songs which might be one other reason for Reggeaton's popularity. 

The visualisations I created showed very interesting results and I want to look further at audio features like danceability, loudness, tempo and speechiness as a perceive those as very characteristic of both music styles. More importantly want to look more deeply into my results and create more sophisticated visualisations.

Additionally, I still want improve the layout of my portfolio and also compare the two genres looking at specific songs. 